---
title: "dataviz-project-YZ"
author: "Yingxin Zhang"
date: "2/8/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load-packages, message = TRUE}
library(tidyverse)
```

# Data load
```{r}
salc = read.table("data/student-mat.csv", sep = ",", header=TRUE)
print(nrow(student_alcohol)) # 382 students
```
# Visulize 1: Weekend Alcohol Consumption V.S. Final Grade for 2 Schools
```{r}
ggplot(data = salc, mapping = aes(x = Walc, y = G3, color = school, shape = school)) +
geom_jitter(width = 0) +
geom_point(alpha = 0.5, size = 2) +
ggtitle("Weekend Alcohol Consumption V.S. Final Grade for 2 Schools")
```
## The x axis is Workday alcohol consumption (numeric: from 1 - very low to 5 - very high). The y axis is the final grade. There are 2 different schools: Gabriel Pereira (GP) and Mousinho da Silveira (MS)


# Visulize 2: Alcohol Consumption with Parents's Job
```{r}
ggplot(data = salc, mapping = aes(x = Mjob, y = Walc, fill=factor(Walc))) +
geom_bar(stat ="identity") +
xlab("Mjob") +
ylab("Count") +
ggtitle("Alcohol Consumption with Mother's Job")

ggplot(data = salc, mapping = aes(x = Fjob, y = Walc, fill=factor(Walc))) +
geom_bar(stat ="identity") +
xlab("Fjob") +
ylab("Count") +
ggtitle("Alcohol Consumption with Father's Job")
```
## Given parents are in different types of job, we can see the distribution of child's alcohol consumption is different.

